---
title: "EDA-challenge"
author: "Karim Primov"
date: "2/6/2022"
output: html_document
---

```{r}
library(sjmisc)
library(tidyverse)
f <- "https://raw.githubusercontent.com/difiore/ada-2022-datasets/main/data-wrangling.csv"
d <- read_csv(f, col_names = TRUE) #creates a tibble
d
```

creating variable named "BSD" [ratio of male to female body mass]
```{r}
d$bsd <- (d$Body_mass_male_mean/d$Body_mass_female_mean)
d$bsd
```
creating variable named "sex_ratio" [ratio of adult females to adult males]
```{r}
d$sex_ratio <- (d$AdultFemale/d$AdultMales)
d$sex_ratio
```
creating variable named "DI" [ratio of day range length to diameter of home range]
```{r}
d$DI <- (d$DayLength_km/d$HomeRange_km2)
d$DI
```
Plotting the relationship between day range length and time spent moving
```{r}
plot(d$DayLength_km~d$Move)
```


It appears that there's not much of a relationship between time spent moving and day range length, however, most species appear to spend less than 40 mins per day moving



Plotting the relationship between day range length and log(time spent moving)
```{r}
plot(d$DayLength_km~log(d$Move))
```


There still appears to not be much of a relationship 


Plotting the relationship between log(day range length) and time spent moving
```{r}
plot(log(d$DayLength_km)~d$Move)
```


Data seems skewed to the left

Plotting the relationship between log(day range length) and log(time spent moving)
```{r}
plot(log(d$DayLength_km)~log(d$Move))
```



Log-transforming both variables shows that overall, species that do spend more time moving travel farther than those who don't



Plotting the relationship between log(day range length) and log(time spent moving), grouped by family
```{r}
p <- ggplot(data = d, aes(x = log(Move), 
                          y = log(DayLength_km),
                          color = factor(d$Family)
                          ))
p <- p+ xlab("log(Move)") + ylab("log(DayLength_km)")
p <- p + geom_point(na.rm = TRUE)
p <- p + theme(legend.position = "bottom", legend.title = element_blank())
p
```



adding marginal univariate plots
```{r}
library(ggExtra)
ggMarginal(p, type = "densigram")
detach(package:ggExtra)
```




setting up a grid for faceting by a grouping variable
```{r}
p <- p + facet_wrap(~Family, ncol = 4)
p <- p + theme(legend.position = "none")
p
```



Plotting the relationship between day range length and time group size
```{r}
plot(d$DayLength_km~d$MeanGroupSize)
```




Plotting the relationship between log(day range length) and time group size
```{r}
plot(log(d$DayLength_km)~d$MeanGroupSize)
```




Plotting the relationship between day range length and log(time group size)
```{r}
plot(log(d$DayLength_km)~d$MeanGroupSize)
```




Plotting the relationship between log(day range length) and log(time group size)
```{r}
plot(log(d$DayLength_km)~log(d$MeanGroupSize))
```




Plotting the relationship between day range length and time group size, grouped by family
```{r}
p <- ggplot(data = d, aes(x = MeanGroupSize, 
                          y = DayLength_km,
                          color = factor(d$Family)
                          ))
p <- p+ xlab("log(MeanGroupSize)") + ylab("log(DayLength_km)")
p <- p + geom_point(na.rm = TRUE)
p <- p + theme(legend.position = "bottom", legend.title = element_blank())
p
```




adding marginal univariate plots
```{r}
library(ggExtra)
ggMarginal(p, type = "densigram")
detach(package:ggExtra)
```





setting up a grid for faceting by a grouping variable
```{r}
p <- p + facet_wrap(~Family, ncol = 4)
p <- p + theme(legend.position = "none")
p
```



no apparent relationship between day range length and mean group size



Plotting the relationship between log(day range length) and log(time group size), grouped by family
```{r}
p <- ggplot(data = d, aes(x = log(MeanGroupSize), 
                          y = log(DayLength_km),
                          color = factor(d$Family)
                          ))
p <- p+ xlab("log(MeanGroupSize)") + ylab("log(DayLength_km)")
p <- p + geom_point(na.rm = TRUE)
p <- p + theme(legend.position = "bottom", legend.title = element_blank())
p
```




adding marginal univariate plots
```{r}
library(ggExtra)
ggMarginal(p, type = "densigram")
detach(package:ggExtra)
```





setting up a grid for faceting by a grouping variable
```{r}
p <- p + facet_wrap(~Family, ncol = 4)
p <- p + theme(legend.position = "none")
p
```



Cebidae appears to have positive relationship between day range length and mean group size, when both variables log-transformed



Plotting the relationship between body size dimorphism and canine size dimorphism, both overall and by family

Plotting the relationship between body size dimorphism and canine size dimorphism, overall
```{r}
plot(d$Canine_Dimorphism~d$bsd)
```




Plotting the relationship between log(body size dimorphism) and log(canine size dimorphism), overall
```{r}
plot(log(d$Canine_Dimorphism)~log(d$bsd))
```



Points more spread out, positive, nonlinear relationship between bsd and Canine dimorphism

Plotting the relationship between body size dimorphism and canine size dimorphism, grouped by family
```{r}
p <- ggplot(data = d, aes(x = bsd, 
                          y = Canine_Dimorphism,
                          color = factor(d$Family)
                          ))
p <- p+ xlab("bsd") + ylab("Canine_Dimorphism")
p <- p + geom_point(na.rm = TRUE)
p <- p + theme(legend.position = "bottom", legend.title = element_blank())
p
```




adding marginal univariate plots
```{r}
library(ggExtra)
ggMarginal(p, type = "densigram")
detach(package:ggExtra)
```




setting up a grid for faceting by a grouping variable
```{r}
p <- p + facet_wrap(~Family, ncol = 4)
p <- p + theme(legend.position = "none")
p
```



Cercopithecidae exhibits somewhat of a linear relationship between bsd and Canine dimorphism, Cebidae does as well yet with fewer data points (limited certainty)



Plotting the relationship between log(body size dimorphism) and log(canine size dimorphism), grouped by family
```{r}
p <- ggplot(data = d, aes(x = log(bsd), 
                          y = log(Canine_Dimorphism),
                          color = factor(d$Family)
                          ))
p <- p+ xlab("log(bsd)") + ylab("log(Canine_Dimorphism)")
p <- p + geom_point(na.rm = TRUE)
p <- p + theme(legend.position = "bottom", legend.title = element_blank())
p
```




adding marginal univariate plots
```{r}
library(ggExtra)
ggMarginal(p, type = "densigram")
detach(package:ggExtra)
```


setting up a grid for faceting by a grouping variable
```{r}
p <- p + facet_wrap(~Family, ncol = 4)
p <- p + theme(legend.position = "none")
p
```




Cercopithecidae also exhibits somewhat of a linear relationship between log (bsd) and log (Canine dimorphism), and Cebidae does as well yet with fewer data points (limited certainty)


Diet strategy analysis 


```{r}
d <- mutate(
  d,
  "diet" = case_when(
    Fruit >= 50 ~ "frugivore", 
    Leaves >= 50 ~ "folivore",
    Fruit < 50 & Leaves < 50 ~ "omnivore"
  )
)

```




Boxplots of group size vs dietary strategies
```{r}
boxplot(d$MeanGroupSize~d$diet)
```




Frugivores, on average, live in smaller groups than folivores, however, there's much larger variability in frugivore data
compared to folivores


One line code using forward pipe operator

```{r}
s <- mutate(d, Binomial = paste(Genus, Species, sep = " ")) %>%
  select(Binomial,
         Family,
         Brain_Size_Species_Mean,
         Body_mass_male_mean
         ) %>%
  group_by(Family) %>%
  summarise(
    avgbrainsize = mean(Brain_Size_Species_Mean, na.rm = TRUE),
    avgbodymass = mean(Body_mass_male_mean, na.rm = TRUE),
  ) %>%
  arrange(desc(avgbrainsize))
s

```



















































